### 2022-09-19 main/chapter_context.tex
merge theory and experiment chapters

\begin{singlespacing}
\chapter{The ATLAS experiment}
\label{chapter:experiment}
%
\begin{epigraphs}
\qitem{%
if observing outer space gives us a view of the past, observing inner space
would surely give us a glimpse into the future --- would be interesting if NASA
made a telescope for that%
}%
{Ken~M,
\textit{Comment: Yahoo! News},
2012~\cite{kenm2012inner}}
\end{epigraphs}
\end{singlespacing}

\TODO{merge into theory chapter?}

@@

\section{Inner tracker}
\label{sec:atlas_tracker}
\TODO{inner tracker}
\begin{displayquote}
``%
The inner tracking detector covers the pseudorapidity range $|\eta| < 2.5$.
It consists of silicon pixel, silicon microstrip, and transition radiation
tracking detectors.
An additional layer of silicon pixels, the insertable
B-layer~\cite{ATLAS-TDR-19, PIX-2018-001}, was installed before Run~2.%
''~\cite{atlas2022searches}
\end{displayquote}

\section{Calorimeters}
\label{sec:atlas_calo}
\TODO{calo}
\begin{displayquote}
``%
Lead/liquid-argon (LAr) sampling calorimeters provide electromagnetic (EM)
energy measurements with high granularity.
A steel/scintillator-tile hadron calorimeter covers the central pseudorapidity
range ($|\eta| < 1.7$).
The endcap and forward regions are instrumented with LAr calorimeters for both
the EM and hadronic energy measurements up to $|\eta| = 4.9$.%
''~\cite{atlas2022searches}
\end{displayquote}

\section{Muon system}
\label{sec:atlas_muon}
\TODO{muon system}
\begin{displayquote}
``%
The muon spectrometer surrounds the calorimeters and is based on three large
superconducting air-core toroidal magnets with eight coils each.
The field integral of the toroids ranges between $2.0$ and
$6.0\,\mathrm{T\kern-0.15ex m}$ across most of the detector.
The muon spectrometer includes a system of precision chambers for tracking and
fast detectors for triggering.%
''~\cite{atlas2022searches}
\end{displayquote}

\section{Trigger}
\label{sec:atlas_trigger}
% L1 L2 HLT
\TODO{trigger}
\begin{displayquote}
``%
A two-level trigger system is used to select events.
The first-level trigger is implemented in hardware and uses a subset of the
detector information to accept events at a rate below $100\,\mathrm{kHz}$.
This is followed by a software-based trigger that reduces the accepted event
rate to $1\,\mathrm{kHz}$ on average depending on the data-taking conditions.%
''~\cite{atlas2022searches}
\end{displayquote}

\section{Software}
\label{sec:atlas_software}
% how do we predict data?
\TODO{software}
\begin{displayquote}
``%
An extensive software suite~\cite{ATL-SOFT-PUB-2021-001} is used in the
reconstruction and analysis of real and simulated data, in detector operations,
and in the trigger and data acquisition systems of the experiment.%
''~\cite{atlas2022searches}
\end{displayquote}

\section{Data}
\label{sec:atlas_data}
% discuss collisions, pileup, met, metsig

>>>>

\begin{singlespacing}
\chapter{Context: theory and experiment}
\label{chapter:context}
%
\begin{epigraphs}
\qitem{%
The idea that theorems follow from the postulates does not correspond to simple
observation.
If the Pythagorean theorem were found to not follow from the postulates, we
would again search for a way to alter the postulates until it was true.
Euclid's postulates came from the Pythagorean theorem, not the other way.}%
{Richard~Hamming,
\textit{The Unreasonable Effectiveness of Mathematics},
1980~\cite{hamming1980unreasonable}}
\qitem{%
One geometry can not be more true than another; it can only be more convenient.
Geometry is not true, it is advantageous.}%
{Robert~M.~Pirsig,
\textit{Zen and the Art of Motorcycle Maintenance},
1974~\cite{pirsig1999zen}}
\end{epigraphs}
\end{singlespacing}

\begin{singlespacing}
\section{The Standard Model and beyond}
%
\begin{epigraphs}
\qitem{%
One geometry can not be more true than another; it can only be more convenient.
Geometry is not true, it is advantageous.}%
{Robert~M.~Pirsig,
\textit{Zen and the Art of Motorcycle Maintenance},
1974~\cite{pirsig1999zen}}
\end{epigraphs}
\end{singlespacing}

\TODO{Standard Model past present future}

\subsection{Particles}

\subsection{The Standard Model}

\subsection{Supersymmetric models}

\subsubsection{Minimal supersymmetric models}
Lightest Supersymmetric Particle (LSP)

\subsubsection{Guage Mediated Supersymmetry Breaking}
Gauge Mediated Supersymmetry Breaking (GMSB)

\begin{singlespacing}
\section{The \atlas\ Experiment}
%
\begin{epigraphs}
\qitem{%
if observing outer space gives us a view of the past, observing inner space
would surely give us a glimpse into the future --- would be interesting if NASA
made a telescope for that%
}%
{Ken~M,
\textit{Comment: Yahoo! News},
2012~\cite{kenm2012inner}}
\end{epigraphs}
\end{singlespacing}

@@

\subsection{Inner tracker}
\label{sec:atlas_tracker}
\TODO{inner tracker}
\begin{displayquote}
``%
The inner tracking detector covers the pseudorapidity range $|\eta| < 2.5$.
It consists of silicon pixel, silicon microstrip, and transition radiation
tracking detectors.
An additional layer of silicon pixels, the insertable
B-layer~\cite{ATLAS-TDR-19, PIX-2018-001}, was installed before Run~2.%
''~\cite{atlas2022searches}
\end{displayquote}

\subsection{Calorimeters}
\label{sec:atlas_calo}
\TODO{calo}
\begin{displayquote}
``%
Lead/liquid-argon (LAr) sampling calorimeters provide electromagnetic (EM)
energy measurements with high granularity.
A steel/scintillator-tile hadron calorimeter covers the central pseudorapidity
range ($|\eta| < 1.7$).
The endcap and forward regions are instrumented with LAr calorimeters for both
the EM and hadronic energy measurements up to $|\eta| = 4.9$.%
''~\cite{atlas2022searches}
\end{displayquote}

\subsection{Muon system}
\label{sec:atlas_muon}
\TODO{muon system}
\begin{displayquote}
``%
The muon spectrometer surrounds the calorimeters and is based on three large
superconducting air-core toroidal magnets with eight coils each.
The field integral of the toroids ranges between $2.0$ and
$6.0\,\mathrm{T\kern-0.15ex m}$ across most of the detector.
The muon spectrometer includes a system of precision chambers for tracking and
fast detectors for triggering.%
''~\cite{atlas2022searches}
\end{displayquote}

\subsection{Trigger}
\label{sec:atlas_trigger}
% L1 L2 HLT
\TODO{trigger}
\begin{displayquote}
``%
A two-level trigger system is used to select events.
The first-level trigger is implemented in hardware and uses a subset of the
detector information to accept events at a rate below $100\,\mathrm{kHz}$.
This is followed by a software-based trigger that reduces the accepted event
rate to $1\,\mathrm{kHz}$ on average depending on the data-taking conditions.%
''~\cite{atlas2022searches}
\end{displayquote}

\subsection{Software}
\label{sec:atlas_software}
% how do we predict data?
\TODO{software}
\begin{displayquote}
``%
An extensive software suite~\cite{ATL-SOFT-PUB-2021-001} is used in the
reconstruction and analysis of real and simulated data, in detector operations,
and in the trigger and data acquisition systems of the experiment.%
''~\cite{atlas2022searches}
\end{displayquote}

\subsection{Data}
\label{sec:atlas_data}
% discuss collisions, pileup, met, metsig

### 2022-09-19 main/chapter_searches.tex
merge theory and experiment chapters

\subsection{Probability, abstract}
A probability $\prob{b}{a}$ is a number, which quantifies the proportion
of $b$ within $a$ and can be expressed as a ratio~\cite{axioms1010038}
\begin{equation}
\label{eqn:searches_prob_ratio}
\prob{b}{a} = \frac{m(b, a)}{m(a)},
\end{equation}
in which $m(x)$ is a measure (a non-negative quantity assigned to its
argument, such as mass of tea in a cup), and the comma `,' means `and'
(logical and or set intersection or any other equivalent operation).
For example, $\prob{b}{a}$ could quantify the proportion of tea within
a spoon within a cup --- then $m(a)$ measures the content of the cup, and
$m(b, a)$ measures the amount of tea in the teaspoon that is partially
submerged in the cup.

@@

\subsection{Probability, applied}
\label{sec:searches_probability_applied}
Science involves making predictions, testing those predictions against data,
and reporting what is learned from the results.
Various probabilities are involved in this prediction and learning, and those
probabilities have conventional names that communicate our intended use for
them.
Most important for prediction is a \textbf{likelihood}
\begin{align}
\label{eqn:searches_likelihood}
L(a) &= \prob{b}{a}
\intertext{%
which explores the probabilities assigned to a fixed result $b$
(usually data) from different contexts $a$ (hypotheses).
A likelihood is clearly a probability, but it is not a probability distribution
--- only the left hand argument $b$ forms a normalized distribution.
Alternatively, fixing the context gives a \textbf{prior}%
}
\pi(b) &= \prob{b}{a}
\label{eqn:searches_prior}
\end{align}
normalized probability distribution over results $b$ for a fixed context $a$.

>>>>

\subsection{Probability: abstract}
A probability $\prob{b}{a}$ is a number, which quantifies the proportion
of $b$ within $a$ and can be expressed as a ratio~\cite{axioms1010038}
\begin{equation}
\label{eqn:searches_prob_ratio}
\prob{b}{a} = \frac{m(b, a)}{m(a)},
\end{equation}
in which $m(x)$ is a measure (a non-negative quantity assigned to its
argument, such as mass of tea in a cup), and the comma `,' means `and'
(logical and or set intersection or any other equivalent operation).
For example, $\prob{b}{a}$ could quantify the proportion of tea within
a spoon within a cup --- then $m(a)$ measures the content of the cup, and
$m(b, a)$ measures the amount of tea in the teaspoon that is partially
submerged in the cup.

@@

\subsection{Probability: applied}
\label{sec:searches_probability_applied}
Science involves making predictions, testing those predictions against data,
and reporting what is learned from the results.
Various probabilities are involved in this prediction and learning, and those
probabilities have conventional names that communicate our intended use for
them.
Most important for prediction is a \textbf{likelihood}
\begin{align}
\label{eqn:searches_likelihood}
L(a) &= \prob{b}{a}
\intertext{%
which explores the probabilities assigned to a fixed result $b$
(usually data) from different contexts $a$ (hypotheses).
A likelihood is clearly a probability, but it is not a probability distribution
--- only the left hand argument $b$ forms a normalized distribution.
Alternatively, fixing the context gives a \textbf{prior}%
}
\pi(b) &= \prob{b}{a}
\label{eqn:searches_prior}
\end{align}
normalized probability distribution over results $b$ for a fixed context $a$.
